---
title: "Assignment 2 - R Markdown"
author: "M. Yeung"
date: "November 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
Read the CSV file into activity

```{r read file}
data <- read.csv('./data/activity.csv')
activity <- na.omit(data)
```


##What is mean total number of steps taken per day?
Plot Histogram, find Mean and Median
```{r historgram}
StepsTotal <- tapply(activity$steps, activity$date, sum)
hist(StepsTotal, main="Histogram of steps taken", xlab="Total steps taken per day")

orgmean <- mean(StepsTotal, na.rm = TRUE)
print(orgmean)
orgmedian <-median(StepsTotal,na.rm=TRUE)
print(orgmedian)
```

##What is the average daily activity pattern?
Average Steps activity per 5 min interval
```{r Average Interval}
StepsInterval <- tapply(activity$steps, activity$interval,  mean, na.rm=TRUE)
plot(StepsInterval, type="l", main="Average steps per interval", xlab="Average number of steps", ylab="Average across all days")
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
names(which.max(StepsInterval))
```

##Imputing missing values
Average Steps activity per 5 min interval
```{r Imputing}
#Total NA
sum(is.na(data$steps))
#fill in missing NA data with Average of that interval
AverageStep <- aggregate(steps ~ interval, data = data, FUN = mean)
newdata <- data
for (i in 1:nrow(data)) {
    if (is.na(newdata[i, ]$steps)) {
        newdata[i,]$steps <- subset(AverageStep, interval == data[i, ]$interval)$steps
    }
}
newStepsTotal <- tapply(newdata$steps, newdata$date, sum, na.rm=TRUE)
hist(newStepsTotal, main="New histogram of steps taken", xlab="Total steps taken per day")
#Mean and Median
newmean<-mean(newStepsTotal, na.rm = TRUE)
print(newmean)
newmedian<-median(newStepsTotal,na.rm=TRUE)
print(newmedian)

#comparing old vs new mean
abs(newmean-orgmean)
#comparing old vs new median
abs(newmedian-orgmedian)

```

##Are there differences in activity patterns between weekdays and weekends?
```{r Week}
library(lattice)
library(plyr)
#Weekday and Weekend variable in new column
newdata$weekdays <- weekdays(as.Date(newdata$date))
newdata$weekdays <- ifelse(newdata$weekdays %in% c("Saturday", "Sunday"),"Weekend", "Weekday")
newStepsInterval <- ddply(newdata, .(interval, weekdays), summarise, steps=mean(steps))
xyplot(steps ~ interval | weekdays, data = newStepsInterval, layout = c(1, 2), type="l", xlab = "5 min interval", ylab = "Total number of steps")
```

